The imputation methods big test

### 2020-04-30 pd2.Rmd
wstępne poprawki

The eight data frames on which we will test these above were taken from **OpenML100 collection** and were corrected specifically for this research. They can be found under the following identifiers with the following names:

@@

In order to read all the eight datasets we will use other function, which will return a dataframe containing all the important informations about the test matrices.

@@

@@

The results of our work will be checked using the implemented classification tree algorithm from the *rpart* package. We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a good model. The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpast should do a good job.

@@

Judging by the logarithmic scale, no surprise removing rows and median replenishment are definetely the fastest methods, with removing rows being several times faster. Looking at more advanced ones, definetely *VIM* rules - probably usually something like 5-10 times faster than *mice* and *missForest*. These last two are quite slow, with the missforest appearing to be slightly faster.

@@

Taking into acount all the imputations which have been implemented, we can clearly see that two easy ones really stand out. However, we cannot fully compare the other three - due to the fact that missForest failed on three data sets, and mice - up to seven.

@@

Let's count these four measures again, but only for those datasats for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).

@@

Now we have better results to compare. Taking into account the mean time, definetely *missForest* is the slowest, but also its standard deviaton is incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets. As shown in the chart, *VIM* is definitely better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so you can expect that for small data sets there is not much difference between them, and a lot of time we definitely need to devote to these larger data frames.

@@

The powerful but also fast tree classifier from *rpast* (both looking at the median and average) counted everything in less than a second. The largest fluctuations can be observed for the method of *median* and *VIM* supplementation - for them the standard deviation is less than two and a half seconds. You can also see very similar all three average measures for these two, probably due to the operation of *VIM* as a result giving similar effects as the *median*. Interestingly, unbeatably the fastest tree was built for the *mice* algorithm ... but wait! Maybe we'd better compare only the imputations for those sets for which we've done everything.

@@

library(gridExtra)
grid.arrange(accPlot +
               ggtitle("Accuracies") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
              prePlot +
               ggtitle("Precisions") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
              recPlot +
               ggtitle("Recalls") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
             f1Plot +
               ggtitle("F1s") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
             nrow = 2,
             top = "Measues values for different datasets' imputations")
```

@@

Visualizations often help and ever give answers to the questions, but this time the best option will be definetely to calculate some measures. What means, medians and standard deviations give these measures?

@@

Let's visualize dataset's model result's measures.

>>>>

The fourteen data frames on which we will test these above were taken from **OpenML100 collection** and were corrected specifically for this research. They can be found under the following identifiers with the following names:

@@

In order to read all datasets we will use other function, which will return a dataframe containing all the important informations about the test matrices.

@@

All of datasets contain missing information denoted as NA and differ in size, number of features, number of observations and percentage of missing data. Table below decribes those informations.

```{r}
### TU KOD HANI BEDZIE DO WKLEJENIA

dfs <- read_all_datasets()
```

@@

The results of our work will be checked using the implemented classification tree algorithm from the *rpart* package. We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a good model. The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpart should do a good job.

@@

Judging by the logarithmic scale, no surprise removing rows and median replenishment are definitely the fastest methods, with removing rows being several times faster. Looking at more advanced ones, definitely *VIM* rules - probably usually something like 5-10 times faster than *mice* and *missForest*. These last two are quite slow, with the missforest appearing to be slightly faster.

@@

Taking into account all the imputations which have been implemented, we can clearly see that two easy ones really stand out. However, we cannot fully compare the other three - due to the fact that missForest failed on three data sets, and mice - up to seven.

@@

Let's count these four measures again, but only for those datasets for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).

@@

Now we have better results to compare. Taking into account the mean time, definitely *missForest* is the slowest, but also its standard deviation is incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets. As shown in the chart, *VIM* is definitely better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so you can expect that for small data sets there is not much difference between them, and a lot of time we definitely need to devote to these larger data frames.

@@

The powerful but also fast tree classifier from *rpart* (both looking at the median and average) counted everything in less than a second. The largest fluctuations can be observed for the method of *median* and *VIM* supplementation - for them the standard deviation is less than two and a half seconds. You can also see very similar all three average measures for these two, probably due to the operation of *VIM* as a result giving similar effects as the *median*. Interestingly, unbeatably the fastest tree was built for the *mice* algorithm ... but wait! Maybe we'd better compare only the imputations for those sets for which we've done everything.

@@

library(gridExtra)
grid.arrange(accPlot +
               ggtitle("Accuracies") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
              prePlot +
               ggtitle("Precisions") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
              recPlot +
               ggtitle("Recalls") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
             f1Plot +
               ggtitle("F1s") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
             nrow = 2,
             top = "Measures values for different datasets' imputations")
```

@@

Visualizations often help and ever give answers to the questions, but this time the best option will be definitely to calculate some measures. What means, medians and standard deviations give these measures?

@@

Let's visualize dataset's model results measures.
